\chapter{Introduction}
\label{ch:introduction}


\section{About Fantasy D100}

Fantasy D100 is a fantasy RPG game system. It is essentially my take of the great OpenQuest rules (based on the SRD version).

This has much more than my custom rules though. It has several additions and (what I believe are) cleaner mechanics that abstracts any extraordinary or supernatural abilities (magic, religions, psionics, etc.) into character `Powers'. At the same time it stays true to OQ's rational of being simple, flexible but still relatively realistic.

Fantasy D100 system is pretty open, both at character generation and during character advancement, in that they don’t tie a character down to a predestined path of skill and powers. Everything is a potential choice that a Player can make during play.

Character generation produces characters that have skills in all the basic areas of expertise, a couple of speciality advanced skills, some powers (e.g. magic) and some skill in at least one or two weapons. Most Fantasy D100 characters start out being able to do most things, a skill area or two that they excel at, have a decent chance in a fight and even have some powers to even out the odds.

Because Fantasy D100 characters start off more rounded there is less of an issue about getting the right mix of skills for the group so it can survive the adventure.

Part I of this book goes through a basic ruleset without concetrating in the intricasies or philosophy of any particular campaign world. It represents everything a Gamemaster needs for a mundane medieval fantasy game (i.e. without magic, fantasy races etc.). 


\section{Disciplines and Powers}
The Disciplines chapters in part two contain several different extraordinary and supernatural options to embellish your campaign. Each chapter is self-contained including everything needed with regard to this discipline. 

A Discipline embodies a group of extraordinary or supernatural abilities called Powers. A character might be knowledgeable to several disciplines. Disciplines can vary from special combat techniques that require extensive training to the ability to channel mystical energies. Some Disciplines are:
\begin{rpg-list}
\item Battle: the character possesses extraordinary battle techniques.
\item Magic: the character possesses magic powers.
\item Arcane Magic: the character possesses magic's deeper secrets.
\item Divine Magic: the character can channel power directly from the Gods.
\item Shamanism: the character in tune with the Spirit World.
	%Psionic: Talents related to raw personal power.
\end{rpg-list}

Which disciplines are available depends on the world that the game is being played in. Fantasy D100 provides a flexible system were adding or removing a Discipline is seamless to the rest of the rules. Each Discipline has its own specific rules and an appropriate skill that represents the character's knowledge and ability in the respective discipline which applies to all Powers of that Discipline.

Each Gamemaster can choose to use only the Disciplines appropriate to their campaign's philosophy.


\section{Gamemaster}
The third part of this book contains rules for the organizer of the game. The Gamemaster has a wide variaty of options to include in the campaign world. One such option is to include fantasy races, like Elves, Dwarves, etc. All the appropriate statistics and abilities with regard to those races will be included in the appropriate chapter.


