\chapter{Battle}
\label{ch:battle}

The Battle Discipline represents years of advanced traning in combat techniques. It enables warriors to reach an understanding of the flow of battle to such a degree that he can gain advantages over their opponents. Some times that is forcing an opponent to leave an opening or positioning themselves in such a way to avoid harm. Other times it enables unprecedented focus and calmness in the midst of battle enabling precise and calculated blows.


\section{Spot Rules}

\subsection{Learning Battle}
Battle is treated as a skill. The base chance for Battle is POWx4. Techniques are learnt separately, but the Battle skill determines the success for applying successfully all Techniques. Characters learn Battle from other characters who know the practice. It costs five Improvement Points to get access to the Battle discipline.

\subsection{Learning Techniques}
Characters learn Battle Techniques from other characters who know the appropriate techniques. Learning techniques costs one Improvement Point per Magnitude point (maximum is 2 Magnitude for all techniques). If a character knows a technique at a lower Magnitude, they only have to pay the difference in Improvement Points to gain the technique at a higher Magnitude.

The maximum number of Battle Techniques a character can learn is INT/3.

\begin{rpg-examplebox}
Bors already knows Weapon Mastery at 1 Magnitude. He wants to learn Weapon Mastery 2, so he must spend only one Improvement Point to gain the technique at that Magnitude.
\end{rpg-examplebox}

Battle Techniques can be learnt from a number of sources. The most typical is from military academies but it can also be from the lone retired veteran. 

In each case the player character must be in good standing with the teacher before they will teach them a technique. If the teacher is indifferent to the player character to start with then they will first need to undertake some kind of service, which can be the focus of a Quest.

\subsection{Using Techniques}
A character can use a technique freely as part of a combat action or reaction. Only a single Battle Technique can be used per combat round. Applying a Battle Technique in the midst of combat requires a successful Battle skill test declared before any dice is rolled.

The result of the Battle skill test depends on its success:
\begin{description}
	\item[Success:] A number of Power Points are deducted from the warrior’s total, equal to the Magnitude of the technique. The technique then takes effect.
	\item[Failure:] The technique does not take effect and the character loses one Power Point.
	\item[Critical:] The warrior has been able to understand the opponent's move particularly well. The character loses one Power Point less than required for the technique (zero for Magnitude 1techniques).
	\item[Fumble:] The warrior misunderstands the intent of their opponent. Rather than losing a single Power Point for failing to use the technique, the warrior loses a number of Power Points equal to its Magnitude. 
\end{description}


\subsection{Techniques Traits}
Unless otherwise stated all Techniques have the following traits.

\begin{rpg-list}
\item They can have only two levels of Magnitude; either one of two.
\item All techniques, unless noted, are instant and apply only for the action or reaction that are rolled for.
\item No technique can be used in combination with special attacks like Great Attack, Disarm, etc.
\end{rpg-list}

Other traits used by techniques are detailed below. 
\begin{description}
	\item[Magnitude (X):] The strength and power of the technique. Also the minimum number of Power Points required to cast it. 
	\item[Non-Variable:] The technique may only be used at the stated Magnitude.
\end{description}

\section{Techniques}

\begin{table*}
\begin{center}
\caption{Battle Techniques}
\label{tab:battle-techniques}
\begin{rpg-table}[|l|c|X|]
        \hline
	\textbf{Technique} & \textbf{Traits} & \textbf{Effect}\\
	Confuse & - &  Can only be used for an attack action and forces the opponent's parry or dodge to be at -10\% per Magnitude.\\
	Deadly Aim & - &  Allows a warrior to make a ranged attack with a +10\% and +1 damage per Magnitude.\\
	Extra Attack & Magnitude 2, Non-Variable &  Allows a warrior to make an additional single close combat or unarmed attack action when in need.\\
	Extra Missile & Magnitude 2, Non-Variable &  Allows a warrior to make an additional single ranged attack action at -20\% when in need.\\
	Extra Parry & Magnitude 2, Non-Variable &  Allows a warrior to make an additional single parry reaction when in need.\\
	Extra Reflex & Magnitude 2, Non-Variable &  Allows a warrior to make an additional dodge reaction when in need.\\
	Impr. Armor & - &  The warrior's effective armor is increased by one per Magnitude.\\
	Impr. Disarm & Magnitude 2, Non-Variable &  The attack is similar to the Disarm special attack but without the penalty.\\
	Impr. Intimidate & - &  The warrior's intimidate is more effective by 10\% per Magnitud.\\
	Impr. Knock-Back & Magnitude 2, Non-Variable &  The attack is similar to the Knock-Back special attack but without the penalty.\\
	Weapon Mastery & - &  Allows a warrior to make a close combat or unarmed attack with a +10\% and +1 damage per Magnitude.\\
        \hline
\end{rpg-table}
\end{center}
\end{table*}

